---
title: "Covid91 vaccine study Final2023"
author: "Sakana Uthayasegar"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H(O)$ : The vaccine is not effective in preventing the spread of Covid.

$H(A)$ : The vaccine is effective in preventing the spread of Covid.

## Males


## Introduction

$H(O)$ : The vaccine is not effective in preventing COVID-19 in males.

$H(A)$ : The vaccine is effective in treating COVID-19 in males.

The following data is used as a clinical test. The finalized data was provided by Proffessor Brownlow. 

## Methods

The response methods that is used in this is Cat~Cat. Following data is based off from the final data provided. The numerical results, bar chart will be refereed to as well. 


##Results

```{r}
maledata = filter (FinalData, sex == "M")
```

## Graphical results

```{r}
barchartGC(~infected+treatment,data=maledata)
```
Placebo and Vaccine are compared in growth. 


```{r}
barchartGC(~infected+treatment, data=,maledata, type="percent")
```
According to the graph, from the males who had COVID-19, statistics show that 30% of them who were vaccinated had symptoms of covid and approximately 70 percent of the placebo males had symptoms as well. Graph provides evidence of the slight difference in males compared between vaccinated and non vaccinated individuals. There was almost no different between them. 

## Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=maledata)
rowPerc(table1)
```

```{r}
colPerc(table1)
```
The numerical data provides statistics of the males who took the vaccine and infected by covid-19. The males who took placebo had approximetely 70 percent chance of no infection and 30 percent lesser chance as a male who got the vaccine. There was little to no difference on males felt no symptoms and got the vaccine. 

## Inferencial results

```{r}
chisq.test(table1)
```
```{r}
chisqtestGC(table1)
```
```{r}
fisher.test(table1)
```
According to the chisq-test, the p value is determined to be a small amount, therefore the null hypothesis is rejected. The p value is 0.0000000000001098.

## Discussion and Conclusion

The tests performed brings to the conclusion that the vaccine is not very effective in treating COVID-19 in male patients. The p value of 0.0000000000001098 is such a small amount that the null hypothesis is rejected. The p value from the fisher test shows to be 0.00000000000004613. Therefore this shows that there is a difference in covid-19 effectiveness in males. 





## Females
$H(O)$ : The vaccine is not effective in preventing COVID-19 in females.

$H(A)$ : The vaccine is effective in treating COVID-19 in females.

The following data is used as a clinical test. The finalized data was provided by Proffessor Brownlow. 

## Methods

The response methods that is used in this is Cat~Cat. Following data is based off from the final data given. The numerical results, bar chart will be refereed to as well. 


##Results

```{r}
femaledata = filter (FinalData, sex == "F")
```

## Graphical results

```{r}
barchartGC(~infected+treatment,data=femaledata)
```
```{r}
barchartGC(~infected+treatment, data=,femaledata, type="percent")
```
According to the graph, from the females who had COVID-19, statistics show that 40% of them who were vaccinated had symptoms of covid and approximately 60 perecent of the placebo females had symptoms as well. Graph provides evidence of the slight difference in females compared between vaccinated and non vaccinated individuals. There was almost no different between them. 

## Numerical results

```{r}
table2 <- xtabs(~infected + treatment, data=femaledata)
rowPerc(table2)
```
```{r}
colPerc(table2)
```
The following numerical data provides the females that got Covid-19. From the data of the ones not vaccinated, 60 percent of the females felt symptoms, compared to the 40% females who were vaccinated who felt symptoms. Based on the data, the females from the placebo group, only 2.04 percent of them were infected in covid-19, while 1.37 percent of the females who were vaccinated were infected with covid 19.

## Inferencial results

```{r}
chisq.test(table2)
```
```{r}
chisqtestGC(table2)
```
```{r}
fisher.test(table2)
```

According to the chisq-test, the p value is determined to be a small amount, therefore the null hypothesis is rejected. The p value is 0.0002732

## Discussion and Conclusion

The tests performed brings to the conclusion that the vaccine is not very effective in treating COVID-19 in female patients. The p value of 0.0002732 is such a small amount that the null hypothesis is rejected. The p value from the fisher test shows to be 0.0002162 Therefore this shows that there is a difference in covid-19 effectiveness in females. 

## LGBTQ

## Introduction

$H(O)$ : The vaccine is not effective in preventing COVID-19 in LGBTQ.

$H(A)$ : The vaccine is effective in treating COVID-19 in LGBTQ.

The following data is used as a clinical test. The finalized data was provided by Proffessor Brownlow. 

## Methods

The response methods that is used in this is Cat~Cat. Following data is based off from the final data provided. The numerical results, bar chart will be refereed to as well. 


##Results

```{r}
LGBTQdata = filter (FinalData, LGBTQ== "gay")
```

## Graphical results

```{r}
barchartGC(~infected+treatment,data=LGBTQdata)
```
```{r}
barchartGC(~infected+treatment, data=,LGBTQdata, type="percent")
```
According to the graph, from the LGBTQ who had COVID-19, statistics show that 80% of them who were vaccinated had symptoms of covid and approximately 1 percent of the placebo females had symptoms as well. Graph provides evidence that the ones that did not get covid 19 and were not vaccinated or vaccinated did not show any differentiation. 

## Numerical results

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQdata)
rowPerc(table3)
```

```{r}
colPerc(table3)
```
The numerical data provides statistics of the LGBTQ  placebo shows they experienced 15.38% of the symptoms, compared to the vaccinated group who experienced 84.62 percent of the symptoms. Data provides evidence that 0.66 percent of the LGBTQ group that had placebo were infected by covid 19, compared to the ones vaccinated is 3.54 percent.


## Inferencial results

```{r}
chisq.test(table3)
```
```{r}
chisqtestGC(table3)
```
```{r}
fisher.test(table3)
```
According to the chisq-test, the p value is determined to be a small amount, therefore the null hypothesis is rejected. The p value is 0.0000006668.

## Discussion and Conclusion
According to the chisq-test, the p value is determined to be a small amount, therefore the null hypothesis is rejected. The p value is 0.0000006668. This determines that there is an effect between LGBTQ subgroup and covid vaccine. 
Based on the fisher exact test, the odd ratio is approximately 0.18% , which shows the odds are 0.18 lower in a variable group compared to the other group. This shows a difference between both groups, therefore the null hypothesis is rejected. 

## Druggies

## Introduction

$H(O)$ : The vaccine is not effective in preventing COVID-19 in druggies.

$H(A)$ : The vaccine is effective in treating COVID-19 in druggies.

The following data is used as a clinical test. The finalized data was provided by Professor Brownlow. 

## Methods

The response methods that is used in this is Cat~Cat. Following data is based off from the final data provided. The numerical results, bar chart will be refereed to as well. 


##Results

```{r}
DrugUserData = filter (FinalData, DrugUser == "yes")
```

## Graphical results

```{r}
barchartGC(~infected+treatment,data=DrugUserData)
```
```{r}
barchartGC(~infected+treatment, data=,DrugUserData, type="percent")
```
The drup users who got covid 19, approximately 80 percent of the vaccinated experienced symptoms and 15% of placebo group experienced symptoms. There was a little no difference between the drug users group that are vaccinated or not. 


## Numerical results

```{r}
table4 <- xtabs(~infected + treatment, data=DrugUserData)
rowPerc(table4)
```


```{r}
colPerc(table4)
```
The numerical data provides evidence that 11.67% of the placebo drug user group felt the covid 19 symptoms when they were infected by covid-19. 88.33% of vaccinated drug users experienced symptoms. The data shows that 4% of the drug users were infected with covid 19 and 3.06% vaccinated ones were infected with covid 19.

## Inferencial results

```{r}
chisq.test(table4)
```


```{r}
chisqtestGC(table4)
```
```{r}
fisher.test(table4)
```
According to the chisq-test, the p value is determined to be a small amount, therefore the null hypothesis is rejected. The p value is 0.00000002842 .

## Discussion and Conclusion

According to the chisq-test, the p value is determined to be a small amount, therefore the null hypothesis is rejected. The p value is 0.00000002842. The p value from the fisher exact test is 0.000000232, which shows the odd ratio is approximately 0.12 percent. This shows that the odds of this occurring are lower in the one group in comparison to the other group. The 95 percent confidence interval from the fisher exact test supports this. 


# Overall Results and Conclusions

To conclude this study, the various subgroups affects the effectiveness of COVID-19 vaccines. 

The tests performed brings to the conclusion that the vaccine is not very effective in treating COVID-19 in male patients. The p value of 0.0000000000001098 is such a small amount that the null hypothesis is rejected. The p value from the fisher test shows to be 0.00000000000004613. Therefore this shows that there is a difference in covid-19 effectiveness in males. 
  Next, the tests performed brings to the conclusion that the vaccine is not very effective in treating COVID-19 in female patients. The p value of 0.0002732 is such a small amount that the null hypothesis is rejected. The p value from the fisher test shows to be 0.0002162 Therefore this shows that there is a difference in covid-19 effectiveness in the female subgroup. When compared between female and male subgroups, women's percentage was lesser in placebo and infection of COVID-19 by about 0.51 percent and 0.22 percent. Even though the numerical wise, the difference does not determine to be a lot, the female subgroup still felt less symptoms with the COVID-19 vaccine.  
  Another subgroup researched is the LGBTQ subgroup. Based on the tests runned, the COVID-19 vaccine does have an effect in the LGBTQ, but not in a good way. The LGBTQ who recieved the COVID-19 vaccine got it, experience higher symptoms than the ones who got the placebo. The LGBTQ members thaqt did not get COVID nor get vaccinated showed no difference compared to the other. The ones who recieved the placebo had infection rate of 0.66% compared to the ones who recieved the vaccine that has a infection rate of 3.54 percent. This is higher amount of different, therefore null hypothesis is rejected due to it's great effect. 
  Lastly, a good comparison of results to the LGBTQ subgroup is the drug users group. The COVID 19 vaccine did have an effect, however not a positive one in the ones that got the vaccine and were infected by COVID 19 with the infection rate of 3.06 percent. The placebo group had an approximate infection rate of 0.4 percent, showing a higher amount of difference between the two, therefore, rejecting the null hypothesis. 
  To conclude all the subgroups, The LGBTQ subgroup had the worst results, while the female subgroup contracted the best results. Males were pretty similar in data to the females and drug users were similar in data compared to the LGBTQ subgroups. 
 Therefore,this study shows that the COVID-19 vaccine would significantly benefit the males and females more, compared to LGBTQ members and drug users.








